Lexicalized Domain Representations for Neural Machine Translation

### 2019-05-21 2019/emnlp/sparse.tex
checked all

For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions). This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain. We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:de in the EMEA domain). \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture. The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types,
  % (LSTMs vs. conditional GRUs),
  etc.} \fyDone{More on this.}
As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study. The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings. The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer. Its precision on a development set are higher than 95\%.}
\fyDone{Explain how} Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels. Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance. A last contrast is with \texttt{LDR}$_{oracle}^{0.5}$ where we change the distribution of training sentence to increase the number of ECB samples: as a result, we see a small decrease for EMEA and EPPS, and a large boost for ECB. This shows that our technique can be used in conjunction to other well known strategies for performing domain adaptation.

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
\hline
\multicolumn{8}{l}{} \\[-9pt]  
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & 41.3\\
\hline
$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & 26.37 \\
$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & 25.62 \\
$\mathtt{FT} (ECB)$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 &  27.35\\
\hline
$DC$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & 40.71\\
\hline
$LDR_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & 42.28\\
$LDR^{v2}_{oracle}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & 42.44\\
$LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & 42.24 \\
$LDR_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & 42.28\\
$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & - \\
  \hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English $\rightarrow$German} \\
\hline
$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 &36.23 \\
\hline
$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & 27.35\\
$\mathtt{FT} (EPPS)$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & 31.42\\
$\mathtt{FT} (ECB)$  & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & 25\\
\hline
$DC$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & 35.56\\
\hline
$LDR_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & 37.15\\
$LDR^{v2}_{oracle}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & 38.03\\
$LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
$LDR_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 32.2 \\
$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the Transformer systems \label{tab:results-trsf}}
\end{table*}

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 &37.43 \\
\hline
$\mathtt{FT} (EMEA)$  & 72.06$_{47.33}$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & 20 \\
$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & 29.87 \\
$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & 22.23 \\
\hline
$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
 \hline
$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 \\
\hline
$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
$LDR^{v2}_{oracle}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & 31.2\\
\hline
$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & 20.53\\
$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & 18.89 \\
$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & 25.89\\
\hline
$DC$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
\hline
$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & 32.49\\
\hline
$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN systems\label{tab:results-rnn}}
\end{table*}
% \clearfloat

@@

An important practical question concerns the balance between the generic and the domain-specific part of the embeddings. In the limit where the domain specific part is very small, we should recover the performance of the \texttt{Mixed} system; conversely, we expect to see a less effective sharing of data across domains by increasing the domain-specific regions. Table~\ref{tab:embedding-size} reports the result of  a series of experiments for the Transformer architecture (English-French) with varying domain-specific sizes allocating between 4 and 64 cells for domain-specific information, and the complement to 512 for the generic part. The differences are overall quite small in our experimental setting, where the training data is relatively limited and does not require to use a large embedding size. We therefore decided to allocate $8$~cells for the domain specific part. This suggests that we could easily accomodate more domains with the same architecture, and even reserve some regions to handle supplementary incoming training data.  
% \fyTodo{Easily accomodate more domains}
\fyToDone{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}

>>>>

For these test sets, our proposal (\texttt{LDR}$_{oracle}$) is consistently better that the mix-domain strategy, with gains that range from very large (for EMEA and ECB) to unsignificant (for EPPS in most conditions). This means that our architecture is somehow able to compensate for the data unbalance and to raise the performance of the multi-domain system close to the best (fine-tuned) system in each domain. We even observe rare cases where the \texttt{LDR}$_{oracle}$ system outperforms fine-tuning (eg.\ Transformer en:de in the EMEA domain). \texttt{LDR}$_{oracle}$ is also better than Domain Control in three conditions out of four, \texttt{DC} being seemingly a better choice for the RNN than for the Transformer architecture. The comparison with the \texttt{WDCMT} model is less favourable for our RNN systems, which are arguably much simpler.\footnote{The direct comparison with WDCMT is difficult as the implementation differs in many ways from our RNN: a different framework, different cell types,
  % (LSTMs vs. conditional GRUs),
  etc.} \fyDone{More on this.}
As expected, ignoring the domain label yields a drop in performance for the two fallback strategies considered in this study. The first, \texttt{LDR}$_{generic}$, only uses the generic part of the embeddings. The second, \texttt{LDR}$_{pred}$, resorts to automatically predicted domain labels\footnote{Our domain classifier uses a bi-LSTM RNN encoder, followed by a simple softmax layer. Its precision on a development set are higher than 95\%.}
\fyDone{Explain how} Note that this decrease is however quite limited (and hardly significant for \texttt{LDR}$_{pred}$), showing that our architecture is quite robust to noisy labels. Even in the worst case scenario where all domain tags are intently wrong (\texttt{LDR}$_{wrong}$), we see that the generic part still ensures a satisfying level of performance. A last contrast is with \texttt{LDR}$_{oracle}^{0.5}$ where we change the distribution of training sentences to decrease the weight of EPPS data and increase the number of ECB samples. As a result, we see a small decrease for EMEA and EPPS, and a large boost for ECB. This shows that our technique can be used in conjunction to other well known strategies for performing domain adaptation.

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
\hline
\multicolumn{8}{l}{} \\[-9pt]  
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & 41.3\\
\hline
$\mathtt{FT} (EMEA)$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & 26.37 \\
$\mathtt{FT} (EPPS)$  & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & 25.62 \\
$\mathtt{FT} (ECB)$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 &  27.35\\
\hline
$\mathtt{DC}$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & 40.71\\
\hline
$\mathtt{LDR}_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & 42.28\\
$\mathtt{LDR}_{oracle}^{0.5}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & 42.44\\
$\mathtt{LDR}_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & 42.24 \\
$\mathtt{LDR}_{pred}$ & 74.29 $_{49.84}$ & 37.73 & 54.01 & 33.34 & 28.32 & 26 & 42.28\\
$\mathtt{LDR}_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & - & - & - & - \\
  \hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English $\rightarrow$German} \\
\hline
$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 &36.23 \\
\hline
$\mathtt{FT} (EMEA)$ & 68.35$_{42.97}$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & 27.35\\
$\mathtt{FT} (EPPS)$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & 31.42\\
$\mathtt{FT} (ECB)$  & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & 25\\
\hline
$\mathtt{DC}$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & 35.56\\
\hline
$\mathtt{LDR}_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & 37.15\\
$\mathtt{LDR}_{oracle}^{0.5}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & 38.03\\
$\mathtt{LDR}_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & 37.14\\
$\mathtt{LDR}_{pred}$ & 70.89$_{46.12}$ & 26.53 & 68.63 & 21.45 & 16.61 & 19.07 & 32.2 \\
$\mathtt{LDR}_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & - & - & - & - \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the Transformer systems \label{tab:results-trsf}}
\end{table*}

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 &37.43 \\
\hline
$\mathtt{FT} (EMEA)$  & 72.06$_{47.33}$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & 20 \\
$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & 29.87 \\
$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & 22.23 \\
\hline
$\mathtt{DC}$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
 \hline
$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 \\
\hline
$\mathtt{LDR}_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
$\mathtt{LDR}_{oracle}^{0.5}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
$\mathtt{LDR}_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
$\mathtt{LDR}_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
$\mathtt{LDR}_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & 31.2\\
\hline
$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & 20.53\\
$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & 18.89 \\
$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & 25.89\\
\hline
$\mathtt{DC}$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
\hline
$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & 32.49\\
\hline
$\mathtt{LDR}_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
$\mathtt{LDR}_{oracle}^{0.5}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
$\mathtt{LDR}_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
$\mathtt{LDR}_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
$\mathtt{LDR}_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & - \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN systems\label{tab:results-rnn}}
\end{table*}
% \clearfloat

@@

An important practical question concerns the balance between the generic and the domain-specific part of the embeddings. In the limit where the domain specific part is very small, we should recover the performance of the \texttt{Mixed} system; conversely, we expect to see a less effective sharing of data across domains by increasing the domain-specific regions. Table~\ref{tab:embedding-size} reports the result of  a series of experiments for the Transformer architecture (English-French) with varying domain-specific sizes allocating between 4 and 64 cells for domain-specific information, and the complement to 512 for the generic part. The differences are overall quite small in our experimental setting, where the training data is relatively limited and does not require to use a large embedding size. We therefore decided to allocate $8$~cells for the domain specific part. This suggests that we could easily accomodate more domains with the same architecture, and even reserve some regions to handle supplementary incoming training data.  
% \fyTodo{Easily accomodate more domains}
\fyDone{Given the ways embeddings are computed, why not add more domains, and test robustness agains data presentation order ?}
